Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-24 evaluation/subsections/beyond dsp.tex
Updates from Overleaf

\section{Music Structure Analysis (MSA) as downstream task}

While the usefulness of deep audio embeddings can be evaluated in several tasks, we have chosen a core MIR task for its popularity and complexity: Music Structure Analysis (MSA). This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

>>>>

\section{Music Structure Analysis as downstream task}

While the usefulness of deep audio embeddings can be evaluated in several scenarios, we have chosen a core MIR task for its popularity and complexity: Music Structure Analysis (MSA). This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

### 2023-05-24 future work/future work.tex
Updates from Overleaf

chapter{}

hybrid autoencoder architecture + UNet stuff

>>>>

\chapter{Future Work}

hybrid autoencoder architecture + UNet stuff
transforms
stems
different takes

Utilize the dB-scale Mel-spectrum magnitude of audio as input data; are a popular choice for input representation in music-related tasks when applying CNNs, as has been demonstrated in various studies. In addition to their wide usage, it's been recently reported that their frequency-domain summarization, grounded in psycho-acoustics, is efficient and challenging to replicate through solely data-driven methods. \cite{Kim2020OneStrategies}

### 2023-05-24 methods/subsections/machine learning.tex
Updates from Overleaf

Throughout this chapter, we aim to provide a comprehensive understanding of the methodology employed in our research, laying the foundation for the presentation and discussion of the results in the following chapters.

\section{Artificial Neural Networks (ANNs)}
Artificial Neural Networks (ANNs) are computational models inspired by biological neural systems. They use interconnected layers of artificial neurons to process information and perform tasks like pattern recognition, classification, and regression. ANNs adjust the connections, or "weights," between neurons during learning to improve performance. The concept of neural networks emerged in the mid-20th century with the work of Warren McCulloch and Walter Pitts. Still, modern neural networks began with Frank Rosenblatt's perceptron in 1957, a single-layer network able to learn linearly separable patterns.

@@

\section{Deep Learning (DL)}

Deep learning is a subfield of machine learning that employs deep neural networks to learn hierarchical data representations, enabling computational models with multiple processing layers to learn data abstractions across numerous levels. These methods have dramatically advanced the state-of-the-art in various domains, including speech recognition, visual object recognition, object detection, drug discovery, genomics, and more \cite{LeCun2015DeepLearning}. The effectiveness of deep learning arises from its nonlinear modeling capabilities and scalability with large datasets.

A central algorithm in deep learning is backpropagation, which uses the forward and backward pass process to efficiently compute gradients of the loss function concerning the network's parameters. The forward pass involves propagating input data through the network to generate predictions. In contrast, the backward pass calculates the error between predictions and ground truth and subsequently returns it through the network. This procedure updates the network's parameters using gradient descent, minimizing the loss function and enhancing the model's performance.

Deep learning has led to breakthroughs in processing complex, high-dimensional data, including images, video, speech, and audio, primarily due to deep convolutional nets. Simultaneously, recurrent nets have illuminated sequential data analysis such as text and speech. Deep learning discovers intricate structures in large datasets, improving our understanding and processing capabilities across various applications.

>>>>

Throughout this chapter, we furnish an exhaustive understanding of the methodical process employed in our research, thereby setting the stage for the presentation and discussion of the results in subsequent chapters.

Our approach could be broadly categorized as cognitive modeling, as it strives to mimic human cognitive processes and problem-solving through a computerized model. We champion exposure-based learning in music and active engagement with an array of musical styles, genres, techniques, and learning methods to foster well-rounded musical proficiency and efficient performance with novel data in practical applications.

\section{Deep Learning Modeling}

Deep learning (DL) is a subfield of machine learning that employs deep neural networks to learn hierarchical data representations, enabling computational models with multiple processing layers to learn data abstractions across numerous levels. These methods have dramatically advanced the state-of-the-art in various domains, including speech recognition, visual object recognition, object detection, drug discovery, genomics, and more \cite{LeCun2015DeepLearning}. The effectiveness of deep learning arises from its nonlinear modeling capabilities and scalability with large datasets.

A central algorithm in deep learning is backpropagation, which uses the forward and backward pass process to compute gradients of the loss function concerning the network's parameters. The forward pass involves propagating input data through the network to generate predictions. In contrast, the backward pass calculates the error between predictions and ground truth and returns it through the network. This procedure updates the network's parameters minimizing the loss function and enhancing the model's performance.

\subsection{Artificial Neural Networks (ANNs)}
Artificial Neural Networks (ANNs) are computational models inspired by biological neural systems. They use interconnected layers of artificial neurons to process information and perform tasks like pattern recognition, classification, and regression. ANNs adjust the connections or weights between neurons during learning to improve performance. The concept of neural networks emerged in the mid-20th century with the work of Warren McCulloch and Walter Pitts. Still, modern neural networks began with Frank Rosenblatt's perceptron in 1957, a single-layer network able to learn linearly separable patterns.

@@

### 2023-05-24 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

Contrastive Learning of Musical Representations (CLMR) \cite{CLMR2021} aims to learn discriminative and valuable representations of music without relying on explicit labels. By leveraging the structure and content of music and contrasting different augmentations or transformations of the same musical piece against randomly generated pieces, CLMR models can effectively capture the underlying known invariances that are explicitly incorporated into the design of a model to improve its robustness and generalization.

In contrast, the model learns unknown invariances from the data without explicit instruction.

@@

\begin{enumerate}
\item \textbf{Data Augmentation:} Create different augmentations or transformations of the same musical piece as positive pairs and use other randomly selected pieces as antagonistic pairs.
\item \textbf{Contrastive Learning:} Train by minimizing a contrastive loss function, encouraging the model to produce similar representations for positive pairs and dissimilar representations for antagonistic pairs.
\item\textbf{Latent Representation:} Utilize the learned representations for various Music Information Retrieval (MIR) downstream tasks, such as music genre classification or similarity, both in the symbolic and time domains.
\end{enumerate}

By following these steps, CLMR models can learn meaningful and robust representations of music, potentially enhancing their performance in various MIR tasks and reducing the need for labeled data.

\subsection{Encoder: SampleCNN}

@@

\begin{figure}[ht]
\centering
\begin{equation}
y_i = \frac{1}{N} \sum_{j=1}^{N} x_{ij}
\end{equation}
\caption[Average Pooling operation]{\small{Representation of the Average Pooling operation. Here, $y_i$ is the output tensor after the average pooling process. Each element $x_{ij}$ is part of the 2D input tensor. $N$ is the count of elements in the second dimension of the input tensor, while $i$ and $j$ are used to iterate over the first and second dimensions of the tensors, respectively.}}
\end{figure}

@@

A Siamese Network is a deep learning architecture introduced by Bromley and LeCun in the early 1990s \cite{Bromley1993SignatureNetwork}. This architecture is explicitly designed for tasks requiring comparing or assessing similarity between two input instances. Its name is derived from the network structure, which consists of two or more identical subnetworks connected in parallel and joined at the output layer. These subnetworks share the same architecture, weights, and hyperparameters, allowing more efficient memory usage and computational complexity.

\subsubsection{Self-Supervised Learning (SSL)}

Self-supervised learning (SSL) is a subfield of machine learning where models learn to generate representations of data by finding patterns and structures within the data without relying on external labels or annotations. In other words, the model learns by creating its supervision signal from the raw, unlabeled data.\cite{audioselfsupsurvey}

This approach to learning is inspired by the way humans and animals learn from their environment, where a great deal of knowledge is acquired through observation and interaction without explicit instruction. Self-supervised learning is a machine learning process where the model trains itself to learn one part of the input from another part of the input. In this process, the unsupervised problem is transformed into a supervised problem by auto-generating the labels.

SSL's main advantages include reducing dependency on labeled data, which can be expensive and time-consuming to collect and annotate. Also, encourages more generalizable and robust representations of the data since they capture the inherent structure and properties of the data rather than relying on human-provided labels.

Further understanding of the cognitive processes in music can be obtained through Jean Piaget's theory of cognitive development. Piaget suggests that children acquire knowledge through sensory and motor experiences in their early stages, with reasoning gradually evolving toward abstract ideas and deductive logic through the development of schemas—cognitive structures that underlie human knowledge and skill \cite{Huitt2003PiagetsDevelopment}. These schemas evolve through an equilibration mechanism that incorporates new information with pre-existing knowledge, involving processes of assimilation and accommodation \cite{audioselfsupsurvey}.

\subsubsection{Known and unknown invariance}

In pattern recognition, \textit{invariance} refers to the property of a model to be robust to certain transformations or changes in the input data. In other words, an invariant model should produce the same or similar output for different instances of the same object, even if they have undergone some changes.

\begin{definition}[Known invariance]
Known invariances are transformations or changes that the model or algorithm is explicitly designed to be robust against. By incorporating these invariances into the model's design, the algorithm is better equipped to generalize and recognize patterns in the presence of these transformations.
\end{definition}

\begin{definition}[Unknown invariance]

Unknown invariances are transformations the model has not been explicitly designed to handle. However, it still manages to be robust due to its learning capabilities. These invariances can emerge from the data and may not be known a priori. 
\end{definition}

@@

\subsubsection{Positive sample generation}

While useful audio transformations tools \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0}

Given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of the chain.

@@

\begin{enumerate}
\item Calculate the minimum and maximum audio chunk lengths in samples:
2\begin{equation}
l_{min} = t_{min} \times S
\end{equation}
\begin{equation}
l_{max} = t_{max} \times S
\end{equation}
\item Generate random audio chunk lengths $l_1, l_2, \ldots, l_{n-1}$ from the uniform distribution on the interval $[l_{min}, l_{max}]$. Calculate the final audio chunk length as:
\begin{equation}
l_n = L_A - \sum_{i=1}^{n-1} l_i
\end{equation}
where $L_A$ is the length of the anchor signal in samples.
\item Split the anchor signal $A$ into audio chunks $C_1, C_2, \ldots, C_n$ according to the calculated audio chunk lengths.
\item Shuffle the audio chunks randomly to get the permuted slices $C_{\sigma(1)}, C_{\sigma(2)}, \ldots, C_{\sigma(n)}$, where $\sigma$ is a random permutation of indices from $1$ to $n$.
\item Concatenate the shuffled audio chunks to generate the negative signal:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

@@

The anchor-positive distance and the anchor-negative distance are computed as the Euclidean distance between the corresponding samples:

@@

The "hardest" negative index for each anchor-positive pair is determined by maximizing the difference between the anchor-negative and anchor-positive distances:

>>>>

Piaget's theory of cognitive development suggests children acquire knowledge through sensory experiences, gradually developing abstract reasoning and schemas—basic cognitive structures \cite{Huitt2003PiagetsDevelopment}. These schemas evolve by integrating new information through assimilation and accommodation \cite{audioselfsupsurvey}.

In pattern recognition, models are designed to be robust against known invariances—input data transformations—allowing consistent output. Conversely, unknown invariances not considered in the model's design can still be handled due to its learning capacity.

Contrastive Learning of Musical Representations (CLMR) learns valuable, discriminative music representations without explicit labels by contrasting positive augmentations of a musical piece against negative ones \cite{CLMR2021}. CLMR works under a machine learning subset called Self-Supervised Learning (SSL), where models self-learn from unlabeled data, creating their supervisory signal \cite{audioselfsupsurvey}. This mimics human learning from observations and interactions, transforming unsupervised problems into supervised ones by auto-generating labels. The benefits of SSL include reduced reliance on labeled data, making data representation more robust and generalizable.

@@

\begin{enumerate}
\item \textbf{Data Augmentation:} Create different augmentations or transformations of the same musical piece as positive pairs and use other randomly selected pieces as antagonistic pairs.
\item \textbf{Contrastive Learning:} Train by minimizing a contrastive loss function, encouraging the model to produce similar representations for positive pairs and dissimilar representations for antagonistic pairs.
\end{enumerate}

For this task, a Triple Siamese Network is suggested, a model architecture proved efficient for music similarity retrieval tasks \cite{contentmusicsimtriplet2020}, intending to minimize the loss function between an anchor, positive, and negative sample, achieved through online triplet mining \cite{Sikaroudi2020OfflinePatches}. The overarching aim is to train a model in a self-supervised setting, which can effectively distinguish between similar and dissimilar samples while considering their underlying composition and production texture.

A Siamese Network is a deep learning architecture introduced by Bromley and LeCun in the early 1990s \cite{Bromley1993SignatureNetwork}. This architecture is explicitly designed for tasks requiring comparing or assessing similarity between two input instances. It comprises two or more identical subnetworks connected in parallel and joined at the output layer. These subnetworks share the same architecture, weights, and hyperparameters, allowing more efficient memory usage and computational complexity.

\subsubsection{Encoder: SampleCNN}

@@

\begin{equation}
y_i = \frac{1}{N} \sum_{j=1}^{N} x_{ij}
\end{equation}

Representation of the Average Pooling operation. Here, $y_i$ is the output tensor after the average pooling process. Each element $x_{ij}$ is part of the 2D input tensor. $N$ is the count of elements in the second dimension of the input tensor, while $i$ and $j$ are used to iterate over the first and second dimensions of the tensors, respectively.

@@

@@

\subsubsection{Positive sample generation chain}

The positive image in our study should preserve its intelligible content when subjected to temporal transformations, regardless of alterations in sonic qualities. Maintaining the temporal structure and meaningfulness of the content allows it to present musical elements identical or strikingly similar to the original track.

While the MIR community has access to helpful audio transformation tools \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0}, the specific requirements of our experiments necessitated the development of our transformation chain: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

@@

We posit that high-level musical content unfolds over time; therefore, we argue that the temporal structure of the negative images related to our anchor should be disrupted. While maintaining similar sonic qualities, the content should be rendered unintelligible.

\begin{enumerate}
\item Calculate the minimum and maximum audio chunk lengths in samples:
\begin{equation}
l_{min} = t_{min} \times S
\end{equation}
\begin{equation}
l_{max} = t_{max} \times S
\end{equation}
\item Generate random audio chunk lengths $l_1, l_2, \ldots, l_{n-1}$ from the uniform distribution on the interval $[l_{min}, l_{max}]$. Calculate the final audio chunk length as:
\begin{equation}
l_n = L_A - \sum_{i=1}^{n-1} l_i
\end{equation}
where $L_A$ is the length of the anchor signal in samples.
\item Split the anchor signal $A$ into audio chunks $C_1, C_2, \ldots, C_n$ according to the calculated audio chunk lengths.
\item Shuffle the audio chunks randomly to get the permuted slices $C_{\sigma(1)}, C_{\sigma(2)}, \ldots, C_{\sigma(n)}$, where $\sigma$ is a random permutation of indices from $1$ to $n$.
\item Concatenate the shuffled audio chunks to generate the negative signal:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

@@

The anchor-positive and anchor-negative distances are computed as the Euclidean distances between the corresponding samples.

@@

The "hardest" negative index for each anchor-positive pair is determined by maximizing the difference between the anchor-negative and anchor-positive distances \cite{XuanImprovedMining}.
